---
output: github_document
---
<!-- badges: start -->
[![R-CMD-check](https://github.com/ncn-foreigners/blocking/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ncn-foreigners/blocking/actions/workflows/R-CMD-check.yaml)
[![test-coverage](https://github.com/ncn-foreigners/blocking/actions/workflows/test-coverage.yaml/badge.svg)](https://github.com/ncn-foreigners/blocking/actions/workflows/test-coverage.yaml)
[![Codecov test coverage](https://codecov.io/gh/ncn-foreigners/blocking/branch/main/graph/badge.svg)](https://app.codecov.io/gh/ncn-foreigners/blocking?branch=main)
[![CRAN status](https://www.r-pkg.org/badges/version/blocking)](https://CRAN.R-project.org/package=blocking)
[![CRAN downloads](https://cranlogs.r-pkg.org/badges/grand-total/blocking)](https://cran.r-project.org/package=blocking)
[![CRAN downloads](https://cranlogs.r-pkg.org/badges/blocking)](https://cran.r-project.org/package=blocking)
[![Mentioned in Awesome Official Statistics](https://awesome.re/mentioned-badge.svg)](https://github.com/SNStatComp/awesome-official-statistics-software)


<!-- badges: end -->
  
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Overview

## Description

This R package is designed to block records for data deduplication and record linkage (also known as entity resolution) using [approximate nearest neighbor algorithms (ANN)](https://en.wikipedia.org/wiki/Nearest_neighbor_search) and graphs (via the `igraph` package).

It supports the following R packages that bind to specific ANN algorithms:

+ [rnndescent](https://cran.r-project.org/package=rnndescent) (default, very powerful, supports sparse matrices),
+ [RcppHNSW](https://cran.r-project.org/package=RcppHNSW) (powerful but does not support sparse matrices), 
+ [RcppAnnoy](https://cran.r-project.org/package=RcppAnnoy), 
+ [mlpack](https://cran.r-project.org/package=mlpack) (see `mlpack::lsh` and `mlpack::knn`).

The package can be used with the [reclin2](https://cran.r-project.org/package=reclin2) package via the `blocking::pair_ann` function.

## Installation

Install the stable version from CRAN:

```{r, eval=FALSE}
install.packages("blocking") 
```

You can also install the development version from GitHub:

```{r, eval=FALSE}
# install.packages("pak") # uncomment if needed
pak::pkg_install("ncn-foreigners/blocking")
```

## Basic usage

Load packages for the examples:

```{r}
library(blocking)
library(reclin2)
```

Generate simple data with three groups (`df_example`) and reference data (`df_base`):

```{r}
df_example <- data.frame(txt = c(
  "jankowalski",
  "kowalskijan",
  "kowalskimjan",
  "kowaljan",
  "montypython",
  "pythonmonty",
  "cyrkmontypython",
  "monty"
))
df_base <- data.frame(txt = c("montypython", "kowalskijan", "other"))
df_example
df_base
```

Deduplication using the `blocking` function. Output contains information:

+ the method used (`nnd` refers to the NN descent algorithm), 
+ number of blocks created (here 2 blocks),
+ number of columns used for blocking, i.e., how many shingles were created by the `text2vec` package (here 28),
+ reduction ratio, i.e., how large the reduction of comparison pairs is (here 0.5714, which means blocking reduces comparisons by over 57%).

```{r}
blocking_result <- blocking(x = df_example$txt)
blocking_result
```

Table with blocking results contains:
+ row numbers from the original data,
+ block number (integers),
+ distance (from the ANN algorithm).

```{r}
blocking_result$result
```

Deduplication using the `pair_ann` function for integration with the `reclin2` package. Use the pipeline with the `reclin2` package:

```{r}
pair_ann(x = df_example, on = "txt") |>
  compare_pairs(on = "txt", comparators = list(cmp_jarowinkler())) |>
  score_simple("score", on = "txt") |>
  select_threshold("threshold", score = "score", threshold = 0.55) |>
  link(selection = "threshold")
```

Linking records using the same function where `df_base` is the "register" and `df_example` is the reference data:

```{r}
pair_ann(x = df_base, y = df_example, on = "txt", deduplication = FALSE) |>
  compare_pairs(on = "txt", comparators = list(cmp_jarowinkler())) |>
  score_simple("score", on = "txt") |>
  select_threshold("threshold", score = "score", threshold = 0.55) |>
  link(selection = "threshold")
```

## See also

See section `Data Integration (Statistical Matching and Record Linkage)` in [the Official Statistics Task View](https://CRAN.R-project.org/view=OfficialStatistics). 

Packages that allow blocking:

+ [klsh](https://CRAN.R-project.org/package=klsh) -- k-means locality sensitive hashing,
+ [reclin2](https://CRAN.R-project.org/package=reclin2) -- `pair_blocking`, `pair_minsim` functions,
+ [fastLink](https://CRAN.R-project.org/package=fastLink) -- `blockData` function.

Other:

+ [clevr](https://CRAN.R-project.org/package=clevr) -- evaluation of clustering, helper functions,
+ [exchanger](https://github.com/cleanzr/exchanger) -- Bayesian Entity Resolution with Exchangeable Random Partition Priors.

## Funding

Work on this package is supported by the National Science Centre, OPUS 20 grant no. 2020/39/B/HS4/00941.
