---
output: github_document
---
<!-- badges: start -->
[![R-CMD-check](https://github.com/ncn-foreigners/blocking/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ncn-foreigners/blocking/actions/workflows/R-CMD-check.yaml)
[![test-coverage](https://github.com/ncn-foreigners/blocking/actions/workflows/test-coverage.yaml/badge.svg)](https://github.com/ncn-foreigners/blocking/actions/workflows/test-coverage.yaml)
[![Codecov test coverage](https://codecov.io/gh/ncn-foreigners/blocking/branch/main/graph/badge.svg)](https://app.codecov.io/gh/ncn-foreigners/blocking?branch=main)
<!-- badges: end -->
  
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Overview

## Description

An R package that aims to block records for data deduplication and record linkage (a.k.a. entity resolution) based on [approximate nearest neighbours algorithms (ANN)](https://en.wikipedia.org/wiki/Nearest_neighbor_search) and graphs (via the `igraph` package).

Currently supports the following R packages that binds to specific ANN algorithms:

+ [rnndescent](https://cran.r-project.org/package=rnndescent) (default, very powerful, supports sparse matrices),
+ [RcppHNSW](https://cran.r-project.org/package=RcppHNSW) (powerful but does not support sparse matrices), 
+ [RcppAnnoy](https://cran.r-project.org/package=RcppAnnoy), 
+ [mlpack](https://cran.r-project.org/package=RcppAnnoy) (see `mlpack::lsh` and `mlpack::knn`).

The package also supports integration with the [reclin2](https://cran.r-project.org/package=reclin2) package via `blocking::pair_ann` function.

## Funding

Work on this package is supported by the National Science Centre, OPUS 22 grant no. 2020/39/B/HS4/00941.

## Installation

You can install the development version of the `blocking` package from GitHub with:

```{r, eval=FALSE}
# install.packages("remotes") # uncomment if needed
remotes::install_github("ncn-foreigners/blocking")
```

## Basic usage

Load packages for the examples

```{r}
library(blocking)
library(reclin2)
```

Generate simple data with two groups (`df_example`) and reference data (`df_base`).

```{r}
df_example <- data.frame(txt = c(
  "jankowalski",
  "kowalskijan",
  "kowalskimjan",
  "kowaljan",
  "montypython",
  "pythonmonty",
  "cyrkmontypython",
  "monty"
))
df_base <- data.frame(txt = c("montypython", "kowalskijan", "other"))

df_example

df_base
```

Deduplication using `blocking` function. Output contains information about:

  + the method used (where `nnd` which refers to the NN descent algorithm), 
  + number of blocks created (here 2 blocks),
  + number of columns used for blocking, i.e. how many shingles were created by `text2vec` package (here 28),
  + reduction ratio, i.e. how large is the reduction of comparison pairs (here 0.5714 which means blocking reduces comparison by over 57%).

```{r}
blocking_result <- blocking(x = df_example$txt)
## data frame with indices and block 
blocking_result
```

Table with blocking which contains:

+ row numbers from the original data
+ block number (integers),
+ distance (from the ANN algorithm).

```{r}
blocking_result$result
```

Deduplication using the `pair_ann` function for integration with the `reclin2` package. Here I use the pipeline that can be used with the `reclin2` package.

```{r}
pair_ann(x = df_example, on = "txt") |>
  compare_pairs(on = "txt", comparators = list(cmp_jarowinkler())) |>
  score_simple("score", on = "txt") |>
  select_threshold("threshold", score = "score", threshold = 0.55) |>
  link(selection = "threshold")
```
Record linkage using the same function where `df_base` is the "register" and `df_example` is the reference (query data).

```{r}
pair_ann(x = df_base, y = df_example, on = "txt", deduplication = FALSE) |>
  compare_pairs(on = "txt", comparators = list(cmp_jarowinkler())) |>
  score_simple("score", on = "txt") |>
  select_threshold("threshold", score = "score", threshold = 0.55) |>
  link(selection = "threshold")
```

## See also

See section `Data Integration (Statistical Matching and Record Linkage)` in [the Official Statistics Task View](https://cran.r-project.org/web/views/OfficialStatistics.html). 

Packages that allow blocking:

+ [klsh](https://CRAN.R-project.org/package=klsh) -- k-means locality sensitive hashing,
+ [reclin2](https://CRAN.R-project.org/package=reclin2) -- `pair_blocking`, `pari_minsim` functions,
+ [fastLink](https://CRAN.R-project.org/package=fastLink) -- `blockData` function.

Other:

+ [clevr](https://CRAN.R-project.org/package=clevr) -- evaluation of clustering, helper functions.
+ [exchanger](https://github.com/cleanzr/exchanger) -- bayesian Entity Resolution with Exchangeable Random Partition Priors
