---
output: github_document
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Overview

## Description

A small package used to block records for data deduplication and record linkage (entity resolution) based on [approximate nearest neighbours algorithms (ANN)](https://en.wikipedia.org/wiki/Nearest_neighbor_search) and graphs (via `igraph`).

Currently supports the following R packages that binds to specific ANN algorithms

+ [RcppHNSW](https://cran.r-project.org/package=RcppHNSW), 
+ [RcppAnnoy](https://cran.r-project.org/package=RcppAnnoy), 
+ [mlpack](https://cran.r-project.org/package=RcppAnnoy) (see `mlpack::lsh` and `mlpack::knn`).


## Funding

Work on this package is supported by the the National Science Centre, OPUS 22 grant no. 2020/39/B/HS4/00941.

## Installation

You can install the development version of `blocking` from GitHub with:

```{r, eval=FALSE}
# install.packages("remotes") # uncomment if needed
remotes::install_github("ncn-foreigners/blocking")
```

## Basic usage

```{r}
library(blocking)
```

```{r}
df_example <- data.frame(txt = c(
  "jankowalski",
  "kowalskijan",
  "kowalskimjan",
  "kowaljan",
  "montypython",
  "pythonmonty",
  "cyrkmontypython",
  "monty"
))
df_example
```

```{r}
blocking_result <- blocking(x = df_example$txt)
## data frame with indices and block 
blocking_result$result
```

## See also

+ [reclin2](https://CRAN.R-project.org/package=reclin2)
+ [klsh](https://CRAN.R-project.org/package=klsh)
