---
title: "Blocking records for deduplication"
author: "Maciej Beręsewicz"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Blocking records for deduplication}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  comment = "#>"
)
```

# Setup

Read required packages

```{r setup}
library(blocking)
library(reclin2)
library(data.table)
```

Read the `RLdata500` data used in the [RecordLinkage](https://CRAN.R-project.org/package=RecordLinkage) package from the [dblink](https://github.com/cleanzr/dblink) Github repository.

```{r}
df <- fread("https://raw.githubusercontent.com/cleanzr/dblink/dc3dd0daf55f8a303863423817a0f0042b3c275a/examples/RLdata500.csv")
head(df)
```
This dataset contains `r nrow(df)` with `r NROW(unique(df$ent_id))` entities.

# Blocking for deduplication

Now we create a new column that concatenates the information in each row. 

```{r}
df[, id_count :=.N, ent_id] ## how many times given unit occurs
df[is.na(fname_c2), fname_c2:=""]
df[is.na(lname_c2), lname_c2:=""]
df[, bm:=sprintf("%02d", bm)] ## add leading zeros to month
df[, bd:=sprintf("%02d", bd)] ## add leading zeros to month
df[, txt:=tolower(paste0(fname_c1,fname_c2,lname_c1,lname_c2,by,bm,bd))]
head(df)
```

In the next step we use the newly created column in the `blocking` function. If we specify verbose, we get information about the progress.

```{r}
df_blocks <- blocking(x = df$txt, ann = "hnsw", verbose=TRUE)
```

Results are as follows:

+ based in `RcppHNSW` we created 133 blocks,
+ it was based on 429 columns (2 character shingles),
+ we have 46 blocks of 2 elements, 43 blocks of 3 elements, ..., 1 block of 17 elements.

```{r}
df_blocks
```
Structure of the object is as follows:

+ `result` - a data.table with identifiers and block IDs,
+ `method` - the method used,
+ `metrics` - based on the `igraph::compare` methods for comparing graphs (here NULL),
+ `colnames` - column names used for the comparison.

```{r}
str(df_blocks,1)
```
The resulting data.table has three columns:

+ `x` - Reference dataset (i.e. `df') -- this may not contain all units of `df',
+ `y` - query (each row of `df`) -- this will return all units of `df`,
+ `block` -- the block ID.

```{r}
head(df_blocks$result)
```

We add block information to the final dataset.

```{r}
df_block_result <- copy(df_blocks$result[order(y),])
df[, block_id := df_block_result$block]
head(df)
```

Finally, we can check in how many blocks the same entities (`ent_id`) are observed. In our example, all the same entities are in the same blocks. 

```{r}
df[, .(uniq_blocks = uniqueN(block_id)), .(ent_id)][, .N, uniq_blocks]
```

