---
title: "Integration with existing packages"
author: "Maciej Beręsewicz"
execute:
  warning: false
  message: false
lang: en
output: 
    html_vignette:
        df_print: kable
        toc: true
        number_sections: true
        fig_width: 6
        fig_height: 4
vignette: >
  %\VignetteIndexEntry{Integration with existing packages}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Setup

```{r setup}
library(blocking)
library(reclin2)
library(fastLink)
```

# Data

In the example we will use the same dataset as in the *Blocking records for record linkage* vignette.

```{r}
census <- read.csv("https://raw.githubusercontent.com/djvanderlaan/tutorial-reclin-uros2021/main/data/census.csv")
cis <- read.csv("https://raw.githubusercontent.com/djvanderlaan/tutorial-reclin-uros2021/main/data/cis.csv")
setDT(census)
setDT(cis)
census[is.na(dob_day), dob_day := ""]
census[is.na(dob_mon), dob_mon := ""]
census[is.na(dob_year), dob_year := ""]
cis[is.na(dob_day), dob_day := ""]
cis[is.na(dob_mon), dob_mon := ""]
cis[is.na(dob_year), dob_year := ""]
census[, txt:=paste0(pername1, pername2, sex, dob_day, dob_mon, dob_year, enumcap, enumpc)]
cis[, txt:=paste0(pername1, pername2, sex, dob_day, dob_mon, dob_year, enumcap, enumpc)]
census[, x:=1:.N]
cis[, y:=1:.N]
```

# Integration with the `reclin2` package

The package contains function `pair_ann` which aims at integration with `reclin2` package. This function works as follows

```{r}
pair_ann(x = census[1:1000], 
         y = cis[1:1000], 
         on = "txt", 
         deduplication = FALSE)
```

Which provides you information on the total number of pairs. This can be further included in the pipeline of the `reclin2` package.

```{r}
pair_ann(x = census[1:1000], 
         y = cis[1:1000], 
         on = "txt", 
         deduplication = FALSE,
         ann = "hnsw") |>
  compare_pairs(on = "txt", comparators = list(cmp_jarowinkler())) |>
  score_simple("score", on = "txt") |>
  select_threshold("threshold", score = "score", threshold = 0.75) |>
  link(selection = "threshold") |> 
  head()
```

# Usage with `fastLink` package


# Usage with `RecordLinkage` package





